### 2018-05-05 revision/response.tex
responding to reviews

\documentclass[11pt]{article} 
\usepackage[letterpaper, headheight=100pt]{geometry}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=2.5cm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{xspace}
\newcommand{\eg}{\emph{e.g.}\xspace}
\newcommand{\ie}{\emph{i.e.}\xspace}
%\newcommand{\WV}[2]{\textcolor{red}{#1}\footnote{\textcolor{red}{#2}}}
\newcommand{\WV}[2]{\textcolor{red}{#1\footnote{\textcolor{red}{#2}}}}
\newcommand{\RC}[2]{\textcolor{purple}{#1\footnote{\textcolor{purple}{#2}}}}
\newcommand{\reply}{\ensuremath{\hookrightarrow}\xspace}

\newcommand{\CortyRPaper}{\textcolor{purple}{Corty and Valdar 2018+ [r/vqtl]}\xspace}
\newcommand{\CortyReanalysisPaper}{\textcolor{purple}{Corty \emph{et al} 2018+}\xspace}
\newcommand{\CortyMethodsPaper}{\textcolor{purple}{Corty and Valdar 2018+ [BVH]}\xspace}

% Bibliography
\usepackage{natbib} \bibpunct{(}{)}{;}{author-year}{}{,}
\bibliographystyle{genetics}
%\addto{\captionsenglish}{\renewcommand{\refname}{Literature Cited}}
\setlength{\bibsep}{0.0pt}

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}

\newcommand{\alert}[1]{\textcolor{red}{#1}}
\usepackage[left,running,mathlines]{lineno}
\renewcommand\linenumberfont{\color{gray}\normalfont\tiny\sffamily}
\usepackage{listings}

\newcommand{\auassign}[2]{\mbox{}\marginpar{\footnotesize\textbf{#1}\newline \textcolor{red}{#2}}}

\lhead{Manuscript ID\\G3/2018/200159}
\chead{\bfseries Response to Reviewer Comments}
\rhead{\today\\William Valdar}
\title{BVH: Response To Editors/Reviewers}

\newcounter{ReviewerNum}
\setcounter{ReviewerNum}{1}

\newcounter{QuestionNum}[ReviewerNum]
\setcounter{QuestionNum}{0}

\newcounter{EditorNum}
\setcounter{EditorNum}{1}

\newcounter{EdQuestionNum}[EditorNum]
\setcounter{EdQuestionNum}{0}

\newcommand{\ReviewerQuestion}[1]{
  \vspace{5pt}\goodbreak
  \noindent\fbox{Reviewer:} #1
  \normalfont\par
}
\newcommand{\EditorQuestion}[1]{
  \vspace{5pt}\goodbreak
  \noindent\fbox{Editor:} #1
  \normalfont\par
}
\newcommand{\Response}[1]{
  \goodbreak
  \textcolor{blue}{#1}
  \normalfont\par
}

\begin{document}
\reversemarginpar

\section*{Editor Comments}

\subsection*{Multi-manuscript Issues (identical across responses to reviewers)}

\EditorQuestion{
  Now that you have all three manuscripts potentially going forward for publication in G3, they can be more tightly linked.
}
\Response{
  We have updated the manuscripts such that each one refers to the other two in the introduction and discussion.
}

\EditorQuestion{
   For instance, one of the other papers could provide the benchmark data that is required for [the software] manuscript.
   So the software paper can include those files and refer to the other manuscript for details.
}
\Response{
  We have incorporated this recommendation, using the dataset from Kumar et al. 2013 that was used in one of the case studies in the manuscript titled ``Mean-Variance QTL Mapping Identifies Novel QTL for Circadian Activity and Exploratory Behavior in Mice''.
  The time required to analyze this dataset, according to number of computer cores available and number of permutations desired, is described in the ``benchmarks'' section of the software paper and the relevant files are included in its Zenodo repository.
  (These files are now in two places, since they are also in the Zenodo repository of the reanalysis paper.)
}

\subsection*{Issues Specific to this Manuscript}

\EditorQuestion{
  As suggested by Reviewer 1 it will be very useful to have a benchmark study with times/memory requirements as a function of number of markers and sample size.
  This study will give those interested on using the package an understanding of how this package would scale.
  It will be important to include in the discussion also elements related to scalability: is the package suitable for very large-p/large-n? 
}
\Response{
  We have added this information in the ``Performance Benchmarks'' section.
  See detailed response to Reviewer 1 comment 3.
}

\EditorQuestion{
  There is a very rich literature in the study of genetic control in environmental variance in both plant and animal breeding as well as model organisms.
  This literature should be mentioned. This comment may also apply to some of the companion papers.
  [Citation recommendations provided by the editor.]
}
\Response{
  We have added a section in the introduction that describes how projects in plant and animal breeding seeking to modulate environmental variance are related to the work presented here.
}

% Mulder HA, Bijma P, Hill WG. Selection for Uniformity in Livestock by Exploiting Genetic Heterogeneity of Residual Variance. Genet Sel Evol. 2008;40:37-59.

% Hill WG, Zhang XS. Effects on Phenotypic Variability of Directional Selection Arising Through Genetic Differences in Residual Variability. Genet Res (Camb). 2004;83:121-132.

% Yang Y, Schön CC, Sorensen D. The genetics of environmental variation for dry matter grain yield in maize. Genet Res (Camb). 2012;94:113-119.

% Ibáñez N, Moreno A, Nieto B, Piqueras P, Salgado C, Gutierrez JP. Genetic Parameters Related to Environmental Variability of Weight of Mice; Signs of Correlated Canalised Response. Genet Sel Evol. 2008;40:279-293.

% Ibáñez N, Varona L, Sorensen D, Noguera JL. A study of heterogeneity of environmental variance for slaughter weight in pigs. Animal. 2007;2:19-26.

% Sørensen P, de los Campos G, Morgante F, Mackay TFC, Sorensen D. Genetic Control of Environmental Variation of Two Quantitative Traits of Drosophila melanogaster Revealed by Whole-Genome Sequencing. Genetics. 2015;201(2):487-497. doi:10.1534/genetics.115.180273.

\section*{Reviewer 1 Comments}

\ReviewerQuestion{
  The authors mention ``in almost all traits of interest in human health and disease''.
  The detection of QTL has also been widely used in plant and animal populations.
}
\Response{
  The reviewer is quite correct and we have edited the introduction to reflect the fact that QTL mapping has been applied productively in plant and animal research.
}

\ReviewerQuestion{
  The authors used a set of simulated data with 3 chromosomes and 11 equally spaced markers.
  I would like to see an example with a larger genotype data set.
}
\Response{
  In order to most clearly illustrate the standard workflow associated with package \texttt{vqtl}, the authors have demonstrated its usage on a an unrealistically-small genome.
  The authors are encouraged by the reviewer's enthusiasm to see the package applied to real datasets.
  This manuscript has been submitted as part of a three-paper companionship, where (call this one paper one), paper two contains two real-data reanalyses and paper three contains one real-data reanalysis.
  Thus, the authors contend it is appropriate to maintain this paper having only limited mention of real data, those mentions in the ``Performance Benchmarks'' section.
}

\ReviewerQuestion{
  The CPU time is indicated for the data set of the example.
  I would like to have the CPU time information in a more realistic data set.
  Perhaps authors can use a free data set from the web.
}
\Response{
  Thank you for this recommendation.
  The authors agree that including this information will make the paper more useful to geneticists who are considering using the package -- we have added two sets of benchmarks.
}
\Response{
  First, we demonstrate the performance of the package according to number of computer cores used and number of permutations conducted.
  For this set of benchmarks, we used the intercross from Kumar et. al 2013 which we reanalyzed in \CortyReanalysisPaper, which involved 244 mice, 93 markers, and after using a hidden Markov model to estimate genotype probabilities at locations between markers, 582 loci.
  As recommended by the editor, this example helps tie together the companions, because the results of this analysis are presented in a companion manuscript.
  This addition can be found on page 4.
}
\Response{
  Second, we demonstrate the performance of the package according to the size of the mapping population and the number of genetic markers, holding constant the number of permutations at 1000 and the number of cores at 32.
  For this set of benchmarks, we used crosses simulated with \texttt{R/qtl} so we could control the population size and number of genetic markers.
  This addition can be found on page 4.
}

\ReviewerQuestion{
  In the simulation, the authors should indicate the magnitude of the genetic variance explained by the QTL.
}
\Response{
  Thank you for this recommendation.
  We have added this information to the manuscript and it now appears on page 2.
}

\ReviewerQuestion{
  I think the authors should also indicate if the software is capable in terms of computing power to analyze the data from SNP genotyping devices or even sequencing.
}

\Response{
  Thank you for this recommendation.
  We have addressed this question in the discussion section, on page 4.
}

\end{document}